---
output:
  pdf_document: default
  html_document: default
  word_document: default
---
<center>
  <h1>CS108 STATISTIS</h1>
  <h1>Summer 2019</h1>
  <h1>Assignment 1</h1>
  <h1>Gurgen Hayrapetyan</h1>
</center>

### Load “Trucking_jobs.csv” datain your R environment.

```{r}
jobs <- read.csv("jobs.csv")
```

### How many observations are in the data set?

We can find the number of observation by looking at length of one of the variables

```{r}
observations <- length(jobs$sex)
print(observations)
```

### What are the variables/features and their data types? Indicate if they are categorical or quantitative variables.

To answer this question we can use `str` function

```{r}
str(jobs)
```

From this we can see that `sex` and `title` variables are categorical while `earnings`, `age`, and `hiredyears` are quantative

### What is the mean, standard deviation, median, and range of earnings.

we can access earning from our dataset by `jobs$earnings`.

```{r}
earnings <- jobs$earnings
earnings_mean <- mean(earnings)
earnings_sd <- sqrt(var(earnings))
earnings_median <- median(earnings)
earnings_range <- range(earnings)[2] - range(earnings)[1]

# code for visualizing the values

data <- c(earnings_mean, earnings_sd, earnings_median, earnings_range)
names(data) <- c(" Mean", "Standard Deviation", "Median", "Range")

labelAll <- paste(names(data), ": ", data, "\n", sep="")
cat(labelAll)
```

### Plot a scatterogram of earnings based on hire year

Our Y-axis will be the hireyears X-axis the earnings

```{r}
x <- jobs$hiredyears
y <- jobs$earnings
plot(x, y, main = "Earnings based on years hired", xlab = "Years Hired", ylab = "Earnings")
```

### Plot the histogram of earnings by count and density

We can plot the histgram by just using the `hist` function

```{r}
hist(earnings)
```

### Find quartiles, 80th quantile(percentile), and IQR of the earnings using quantiles function. What is the median

```{r}
quartiles <- quantile(earnings)
quantile_80 <- quantile(earnings, probs = c(0.8))
earnings_iqr <- IQR(earnings)
```
```{r}
print(quartiles)
print(quantile_80)
print(earnings_iqr)
```

### Plot the boxplot of earnings and separate by sex. Find values of outliers.

```{r}
earnings <- jobs$earnings
sex <- jobs$sex

values <- boxplot(earnings ~ sex)

outliers <- values$out
print(outliers)
```

### Plot pie chart for percent male and female in the data set.

```{r}
# seperating our data by gender
all_males <- subset(jobs, sex == "M")
all_females <- subset(jobs, sex == "F")

# getting number of observations in each sample
all_observations <- length(jobs$sex)
males <- round(length(all_males$sex) * 100 / all_observations)
females <- round(length(all_females$sex) * 100 / all_observations)
data <- c(males, females)

# generating labels
names(data) <- c("Male", "Female")
labelAll <- paste(names(data), "\n", data, "%", sep="")

# plotting the chart
pie(data, main="Male/Female Ratio of Jobs", labels = labelAll)
```